---
title: "index"
author: "Yan Wang"
date: "2023-09-20"
output: html_document
---

## Part1: Fun with functions
### 1A: Exponentional transformation
```{r, echo = TRUE}
Exp <- function(x, k) {
  outcome <- 1
  for (i in 1:k) {
  outcome <- outcome + x^i/factorial(i)
  }
  return(outcome)
}
## examples:
Exp(3, 5)
Exp(6, 11)
```

### 1B: Sample mean and sample standard deviation
```{r, echo=TRUE}
sample_mean <- function(x) {
  n <- length(x)
  return(sum(x)/n)
}

sample_sd <- function(x) {
  n <- length(x)
  m <- sample_mean(x)
  variance <- sum((x - m)^2)/(n - 1)
  return(sqrt(variance))
}
## examples:
x <- rnorm(11, mean = 3, sd = 2)
sample_mean(x)
sample_sd(x)
```

### Part1C: Confidence Intervals
```{r, echo = TRUE}
calculate_CI <- function(x, conf) {
  n <- length(x)
  m <- sample_mean(x)
  se <- sample_sd(x) / sqrt(n)
  alpha <- 1 - conf
  degrees_freedom <- n - 1
  t_score <- qt(p = alpha / 2, df = degrees_freedom, lower.tail = FALSE)
  lower_bound <- m - t_score * se
  upper_bound <- m + t_score * se
  ci <- c (lower_bound, upper_bound)
  return(ci)
}
## examples:
data <- 1:1000
x <- sample(data, 16)
calculate_CI(x, 0.95)
calculate_CI(x, 0.99)

```


## Part2: Wrangling data
```{r, echo = FALSE}
tuesdata <- tidytuesdayR::tt_load("2020-01-07")
rainfall <- tuesdata$rainfall
temperature <- tuesdata$temperature
```


```{r}
library(tidyverse)
library(lubridate)
glimpse(rainfall)
glimpse(temperature)
rainfall1 <- drop_na(rainfall)
rainfall2 <- mutate(rainfall1, date = as.Date(ymd(paste(rainfall1$year, rainfall1$month, rainfall1$day, sep = "-"))))
rainfall3 <- select(rainfall2, -month, -day)
rainfall4 <- mutate(rainfall3, city_name = str_to_upper(city_name))
df <- inner_join(x = rainfall4, y = temperature, by = c("city_name", "date"))
print(df)
```

## Part3: Data visualization
### Part 3A: Plotting temperature data over time
```{r, echo = TRUE}
library(ggplot2)
library(remotes)
df1 <- filter(df, year >= 2014)
remotes::install_github("MatthewBJane/theme_park")
ggplot(df1, aes(x = date, y = temperature, color = temp_type)) + geom_point(size = 0.5) + geom_line() + labs(title = "Max/Min Temperature Over Time (2014 and later)", subtitle = "For different cities, the tendency of temperature changing over time seems consistent, showing a wave-like change.", caption = "By Yan Wang", x = "Date", y = "Temperature/°C", color = "Temperature type") + facet_wrap(~city_name) + theme_classic() + theme(plot.background = element_rect(fill = "#FCABCC"), plot.title = element_text(size = 16, face = "bold"), plot.subtitle = element_text(size = 11, face = "italic"), plot.caption = element_text(size = 16, face = "bold"), legend.position = "right", legend.text = element_text(size = 11), legend.title = element_text(size = 13)) + ThemePark::theme_avatar()

```

### Part 3B: Plotting rainfall over time

#### Answer: Considering that not all combinations of city_name and year have measurements, I use *if-else* control structure: when there are valid values in the combination of city_name and year, then to plot the histogram, otherwise, stop the next operation and report an "Error: There is no data about the city in this year!". To finish this idea, I use *!any()* function to help me check out whether there is valid values after inputting "city" and "year". Also, I use *&* to make sure that both conditions are "TRUE" at the same time. About the histogram, I customize it by adding color to bars, adding title, subtitle, caption, x-label, and y-labe, etc. I also use the barbie theme to make it shinny. Specifically, in order to show the special input "city_name" and "year" in subtitle, I also use the *paste()* function and *deparse(substitute())* function.
```{r}
newdf <- data.frame(df)
distribution_plot <- function(city, years) {
  if (!any(newdf$city_name == city & newdf$year == years)) {
  stop("Error: There is no data about the city in this year!")  
  } else {
  new <- filter(newdf, city_name == city, year == years)
  ggplot(new, aes(log(rainfall))) + geom_histogram(binwidth = 0.2, fill = "#87CEEB", color = "white") + labs(title = "The distribution of rainfall", subtitle = paste("For", deparse(substitute(city_name)), "in", deparse(substitute(year))), caption = "By Yan Wang", x = "Log(rainfall)", y = "Frequency") + theme_light() + theme(plot.title = element_text(size = 16, face = "bold"), plot.subtitle = element_text(size = 14, face = "italic"), plot.caption = element_text(size = 16, face = "bold") )+ ThemePark::theme_barbie()
 } 
}

### for example
distribution_plot("PERTH", 1984)
```

## Part4: Apply functions and plot
### Part4A
```{r, echo = TRUE}
rain_df <- data.frame()
cities <- c("BRISBANE", "CANBERRA", "MELBOURNE", "PERTH", "SYDNEY")
years <- c(2014, 2015, 2016, 2017, 2018, 2019)
for (city_name1 in cities) {
   for (year1 in years) {
   df3 <- filter(df1, city_name == city_name1, year == year1)
   x <- df3$rainfall
   m <- sample_mean(x)
   s <- sample_sd(x)
   ci <- calculate_CI(x, 0.95) 
   result <- data.frame(City = city_name1, Year = year1, Mean = m, SD = s, lower_bound = ci[1], upper_bound = ci[2])
   rain_df <- rbind(rain_df, result)
   }
}
print(rain_df)
```
### Part4B
```{r}
ggplot(rain_df, aes(x = Year, y = Mean, color = City)) + geom_point() + geom_line()+ geom_errorbar(aes(ymin = lower_bound, ymax = upper_bound), width = 1, size = 0.2, color = "black", linetype = "dashed") + labs(title = "Mean Rainfall over Years", x = "Years", y = "Mean rainfall(95%CI)", subtitle = "Different cities showing different rainfall tendency over years", caption = "By Yan Wang") + facet_wrap(~City) + theme_classic() + theme(plot.title = element_text(size = 16, face = "bold"), plot.subtitle = element_text(size = 14, face = "italic"), plot.caption = element_text(size = 16, face = "bold") ) + xlim(2013, 2020) + ylim(1, 9) + scale_y_continuous(breaks = seq(1, 9, 0.5), labels = seq(1, 9, 0.5))

```



